---
title: 'Lab 8 Part 1: point pattern analysis'
author: "Grace Bianchi"
date: "2023-03-02"
output: html_document
---

```{r setup, include=TRUE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(spatstat)
library(terra)
library(sf)
library(tidyverse)
library(here)
library(tmap)

```

## Read in red tree vole data

```{r}
# data source name
voles_sf <- read_sf(dsn = here("redtreevoledata"),
                    layer = "ds033") %>% 
  janitor::clean_names() %>% 
  select(county) %>% 
  filter(county == "HUM") %>% 
  st_transform(crs= 32610)

plot(voles_sf)

```

## Get Humboldt County outline

```{r}
humboldt_sf <- read_sf(here("redtreevoledata/california_county_shape_file.shp")) %>% 
  janitor::clean_names() %>% 
  filter(name == "Humboldt") %>% 
  select(name) %>% 
  # coerce lat and long values for county into crs
  st_set_crs(4326) %>% 
  st_transform(32610)


```


## Plot data and county

```{r}
ggplot() +
  geom_sf(data = humboldt_sf, 
          color = "darkorchid",
          fill = "darkorchid4",
          size = 1) + # width of border
  geom_sf(data = voles_sf,
          color = "orange",
          alpha = 0.7,
          size = 2) +
  theme_void()
```

Spatial point pattern analysis

## Convert 'sf' objects to 'spatstat' objects

```{r}
### convert voles observations to spatial point pattern
voles_ppp <- as.ppp(voles_sf) 

### convert county to owin (observation window) object
humboldt_win <- as.owin(humboldt_sf)

### combine together into point pattern object(points + window)
voles_full <- ppp(voles_ppp$x, voles_ppp$y, window = humboldt_win)

plot(voles_full)
```

## Make a kernal density plot

```{r}
# makes zero dimensional point have width
voles_density <- density(voles_full, sigma = 5000) # center of probability field, sigma is sd 
plot(voles_density)
```

## Map it in tmap 

```{r}
### convert desnity object to raster
vole_r <- terra::rast(voles_density) # doesn't keep crs information
crs(vole_r) <- crs(voles_sf)

### plot it
tmap_mode("view")
tm_shape(vole_r) +
  tm_raster(palette = "Reds", legend.show= FALSE)
```


## Let's statistical analysis with a G function (nearest neighbor)

```{r}
### make a sequence of distances over which we'll calculate G(r)
# vector of radiuses
r_vec <- seq(0, 10000, by = 100)

# 
gfunction <- envelope(voles_full, fun = Gest, r = r_vec,
                      nsim = 100)

gfunction_long <- gfunction %>% 
  as.data.frame() %>% 
  pivot_longer(names_to = "model",
               values_to = "g_val",
               cols = obs:hi)

ggplot(data = gfunction_long, aes(x = r, y = g_val)) +
  geom_line(aes(color = model, group = model)) +
  theme_minimal() +
  labs(x = "radius (m)",
       y = "G(r)")
```

Theoretical line is perfect spatial randomness. May be smoother if we had more similations or looked at radius every 10 meters. SInce observations were well above high, we can say that voles are statistically clustered. Percent of voles with neighbors in given radius.


## Let's try an L function

```{r}
r_vec2 <- seq(0, 80000, by = 5000)

lfunction <- spatstat::envelope(voles_full, fun = Lest,
                      r = r_vec2,
                      nsim = 10)

lfunction_long <- lfunction %>% 
  as.data.frame() %>% 
  pivot_longer(cols = obs:hi,
               names_to = "model",
               values_to = "l")

ggplot(data = lfunction_long, aes(x = r, y = l)) +
  geom_line(aes(color = model, group = model)) +
  theme_minimal() +
  labs(x = "radius (m)", y = "L(r)")

```
As we increase radius, observed values are outside the bounds of the theoreical simulations
